---
title: "R Markdown Yahriel Salinas-Reyes HW 2"
author: "Yahriel Salinas-Reyes"
date: "`r Sys.Date()`"
output: word_document
---

```{r, include=FALSE}
## clean R environment 
rm(list = ls(all=TRUE))
graphics.off()
shell("cls")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Include Libraries
library(readxl)
library(ggplot2)
library(here)
library(epiR)
library(MASS)
library(nnet)
library(car)
library(PairedData)
library(pwr)
library(dplyr)
library(FSA) #posthoc

```

# Question 1
## 1. State the Ho, Ha Hypothesis
Null hypothesis (Ho): The dietary supplement has no effect on blood pressure. That is, the mean blood pressure after taking the supplement is the same as before (mu_before=mu_after)


Alternative hypothesis (Ha): The dietary supplement reduces blood pressure. That is, the mean blood pressure after taking the supplement is less than before. (mu_before > mu_after)

## Q1.2. 
```{r, out.width="60%"}
# Blood pressure before taking the supplement
before <- c(130, 125, 140, 135, 138, 132, 137, 142, 129, 136, 133, 139, 140, 150)
# Blood pressure after taking the supplement for 8 weeks
after <- c(122, 118, 133, 127, 129, 124, 130, 135, 121, 128, 126, 132, 120, 123)

# Run a paired t-test
t_test_result <- t.test(after, before, paired = TRUE, alternative = "less")

```

## Q1.3
```{r, out.width="60%"}
# View the results
print(t_test_result)
```
Based on the Output, we get the following results

t-statistic = -6.1753
Degrees of freedom (df) = 13
p-value = .00001674
95% CI for the difference in means: (-INF, -7.030364)

## Q1.4
# If the Ho is true, the probability of having a statistic as extreme as -6.1753 or more extreme, is .001674%.
# Since the value of Ho is outside of the 95% interval # we reject the null hypothesis.
# We conclude that there is a significant reduction in blood pressure after taking the dietary supplement.

# Question 2
```{r, out.width="60%"}

```

# Question 3
```{r, out.width="60%"}

```

# Question 4
```{r, out.width="60%"}

```

# Question 5
```{r, out.width="60%"}

```

# Question 6
```{r, out.width="60%"}

```

# Question 7
```{r, out.width="60%"}

```

# Question 8
```{r, out.width="60%"}

```

# Question 9
```{r, out.width="60%"}

```
























