---
title: "R Markdown Yahriel Salinas-Reyes HW 2"
author: "Yahriel Salinas-Reyes"
date: "`r Sys.Date()`"
output: word_document
---

```{r, include=FALSE}
## clean R environment 
rm(list = ls(all=TRUE))
graphics.off()
shell("cls")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Include Libraries
library(readxl)
library(ggplot2)
library(here)
library(epiR)
library(MASS)
library(nnet)
library(car)
library(PairedData)
library(pwr)
library(dplyr)
library(FSA) #posthoc

```

# Question 1
## Q.1.1 State the Ho, Ha Hypothesis
Null hypothesis (Ho): The dietary supplement has no effect on blood pressure. That is, the mean blood pressure after taking the supplement is the same as before (mu_before=mu_after)


Alternative hypothesis (Ha): The dietary supplement reduces blood pressure. That is, the mean blood pressure after taking the supplement is less than before. (mu_before > mu_after)

## Q.1.2. 
```{r, out.width="60%"}
# Blood pressure before taking the supplement
before <- c(130, 125, 140, 135, 138, 132, 137, 142, 129, 136, 133, 139, 140, 150)
# Blood pressure after taking the supplement for 8 weeks
after <- c(122, 118, 133, 127, 129, 124, 130, 135, 121, 128, 126, 132, 120, 123)

# Run a paired t-test
t_test_result <- t.test(after, before, paired = TRUE, alternative = "less")

```

## Q.1.3
```{r, out.width="60%"}
# View the results
print(t_test_result)
```
Based on the Output, we get the following results

t-statistic = -6.1753
Degrees of freedom (df) = 13
p-value = .00001674
95% CI for the difference in means: (-INF, -7.030364)

## Q.1.4
# If the Ho is true, the probability of having a statistic as extreme as -6.1753 or more extreme, is .001674%.
# Since the value of Ho is outside of the 95% interval # we reject the null hypothesis.
# We conclude that there is a significant reduction in blood pressure after taking the dietary supplement.

# Question 2

## Q.2.1
- Null Hypothesis (Ho): There is no significant difference in productivity before and after the training program (mu_before = mu_after).
- Alternative Hypothesis (Ha): There is a significant difference in productivity before and after the training program (mu_before =/ mu_after)

## Q.2.2
Since we have two sets of related measurements that are paired (before and after), we can use a paired t-test to compare the means. The asumptions for a paired t-test include: 
- The difference between the pairs of data (before-after) should be approximately normally distributed.
We can check this by doing a normality test (Shapiro-Wilk test) and then performing a paired t-test.

```{r, out.width="60%"}
# Data
before <- c(20, 22, 19, 5, 23, 21, 25, 22, 24, 10)
after <- c(15, 26, 21, 10, 27, 23, 60, 26, 27, 50)

# Differences
diff <- before - after

# Check normality of the differences (Shapiro-Wilk test)
shapiro.test(diff)

# Perform a paired t-test
t.test(before, after, paired = TRUE)

```

## Q.2.3
The P-Value = 0.08079

## Q.2.4
Based on the p-value 0.08079, explain the result using the following language 
# If the Ho is true, the probability of having a difference of this magnitude
# or more extreme, is 8.079%
# Since the value of p value 0.08079 is greater than .05 # we fail to reject the null
hypothesis. 
# We conclude that there is not a significant difference in productivity before and after the training program.


# Question 3
```{r, out.width="60%"}
library(ggplot2)
library(dplyr)
library(readxl)
Births <- read_excel("Births.xlsx")
#View(Births)



# Convert GENDER to a factor
Births$`GENDER (1=M)` <- as.factor(Births$`GENDER (1=M)`)
levels(Births$`GENDER (1=M)`) <- c("0", "1")

# Check the structure of the data
str(Births)

# Visualize the birth weight distribution
ggplot(Births, aes(x = `GENDER (1=M)`, y = `BIRTH WEIGHT`, fill = 'GENDER (1=M)')) +
  geom_boxplot() +
  labs(title = "Birth Weight Distribution by Gender",
       x = "Gender (1=M)",
       y = "Birth Weight (g)") +
  theme_minimal()

# Perform normality test for each group (Shapiro-Wilk Test)
shapiro.test(Births$`BIRTH WEIGHT`[Births$`GENDER (1=M)` == "0"])
shapiro.test(Births$`BIRTH WEIGHT`[Births$`GENDER (1=M)` == "1"])

# Since the data may not be normally distributed, use the Mann-Whitney U test
wilcox.test(`BIRTH WEIGHT` ~ `GENDER (1=M)`, data = Births, alternative = "two.sided")

# Display summary statistics
Births %>%
  group_by(`GENDER (1=M)`) %>%
  summarise(median_bw = median(`BIRTH WEIGHT`),
            mean_bw = mean(`BIRTH WEIGHT`),
            sd_bw = sd(`BIRTH WEIGHT`))
```
To test the claim that girls and boys have the same median birth weight using a 0.05 significance level, we can conduct a Mann-Whitney U test (A.K.A. Wilcoxon rank-sum test) since it's a non-parametric test suitable when testing the difference in medians between two independent groups. This test is used instead of a t-test if the assumption of normality is violated.

## Hypothesis
- Null Hypothesis (H0): The median birth weights of boys and girls are the same.
- Alternative Hypothesis (Ha): The median birth weights of boys and girls are different.

## Assumption Checks
- Independence: The birth weight of one child does not influence another child’s birth weight.
- Non-normality: Birth weight data may not be normally distributed. This can be checked visually (using histograms or boxplots) or by a normality test (e.g., Shapiro-Wilk test). If the data is not normal, the Mann-Whitney U test will be appropriate.

## R Output Interpretations
- Boxplot: Visualizes if any obvious difference in birth weights between boys and girls.
- Shapiro-Wilk Test: Checks if birth weights for both genders are normally distributed. If p < 0.05, the data is not normally distributed. FOr both groups, the p-value is less than alpha and therefore not normally distributed. 
- Mann-Whitney U Test: This test outputs a p-value. If p < 0.05, you reject the null hypothesis.

## Conclusion
If p < 0.05: There is sufficient evidence to reject the null hypothesis, meaning there is a significant difference in the median birth weight between boys and girls.
Since the p-value is for both male and female is less than 0.05, we reject the null hypothesis and therefore they are not normally distributed and a significant difference in the median birth weights. 

# Question 4

```{r, out.width="60%"}



```


```{r, out.width="60%"}


```

# Question 5
```{r, out.width="60%"}
# Original data
set.seed(123)
steps <- c(3200, 2900, 3100, 3000, 2800, 3500, 3100, 3000, 3300, 3400,
 2800, 2900, 3200, 3000, 3100, 3300, 3400, 3600, 3500, 3400,
 3000, 2900, 3100, 3200, 3100, 3000, 2800, 2900, 3200, 3100,
 3000, 3200, 3300, 3400, 3500, 3600, 3400, 3300, 3200, 3100)

## Q.5.1
# Plotting histogram of original data
hist(steps, prob = TRUE, main = "Histogram of Original Step Data with Normal Density Curve", xlab = "Steps")
curve(dnorm(x, mean = mean(steps), sd = sd(steps)), add = TRUE, col = "blue", lwd = 2)

## Q.5.2
# Log transformation
log_steps <- log(steps)

# Plotting histogram of log-transformed data
hist(log_steps, prob = TRUE, main = "Histogram of Log-Transformed Step Data with Normal Density Curve", xlab = "Log(Steps)")
curve(dnorm(x, mean = mean(log_steps), sd = sd(log_steps)), add = TRUE, col = "blue", lwd = 2)

## Q.5.3
# Shapiro-Wilk test for original data
shapiro_original <- shapiro.test(steps)

# Shapiro-Wilk test for log-transformed data
shapiro_log <- shapiro.test(log_steps)

# Display results
shapiro_original
shapiro_log

```
## Q.5.4
Interpretation and Conclusion:
- Visual Inspection (Histogram): The histogram of the original data likely shows a skewed distribution, while the log-transformed data may appear closer to normal (depending on how the data looks).
- Normality Test (Shapiro-Wilk): The p-value from the Shapiro-Wilk test helps determine if the data follows a normal distribution. A p-value greater than 0.05 would indicates the data is normally distributed. Comparing the p-values of the original and log-transformed data shows that normality improved.

# Question 6
## Hypothesis
- Null Hypothesis (Ho): The mean arsenic levels in rice from Arkansas, California, and Texas are the same.
- Alternative Hypothesis (Ha): At least one of the states has a different mean arsenic level in rice.

## Assumptions for One-way ANOVA
- Normality: The arsenic levels within each state should be normally distributed. We can use Q-Q plots or the Shapiro-Wilk test to check this.
- Homogeneity of variances: The variances of arsenic levels across the groups should be equal. We can check this using Levene’s test.
- Independent observations: The samples from each state should be independent of each other.
```{r, out.width="60%"}
library(readxl)
Arsenic <- read_excel("Arsenic.xlsx")
View(Arsenic)

Arsenic$STATE <- as.factor(Arsenic$STATE)

# Check assumptions

# 1. Check normality using Q-Q plot for each group
par(mfrow = c(1, 3))  # Set up plotting area
qqnorm(Arsenic$ARSENIC[Arsenic$STATE == "Arkansas"], main = "Arkansas")
qqline(Arsenic$ARSENIC[Arsenic$STATE == "Arkansas"])
qqnorm(Arsenic$ARSENIC[Arsenic$STATE == "California"], main = "California")
qqline(Arsenic$ARSENIC[Arsenic$STATE == "California"])
qqnorm(Arsenic$ARSENIC[Arsenic$STATE == "Texas"], main = "Texas")
qqline(Arsenic$ARSENIC[Arsenic$STATE == "Texas"])

# 2. Perform Levene's test for homogeneity of variance
leveneTest(ARSENIC ~ STATE, data = Arsenic)

# Perform the one-way ANOVA
anova_result <- aov(ARSENIC ~ STATE, data = Arsenic)

# Summary of the ANOVA
summary(anova_result)

# If significant, conduct a post-hoc test to see which groups differ
TukeyHSD(anova_result)

# Visualize the group means
boxplot(ARSENIC ~ STATE, data = Arsenic, main = "Arsenic Levels by State",
        xlab = "State", ylab = "Arsenic Levels", col = c("lightblue", "lightgreen", "lightcoral"))
```
Based on the Post-hoc test, we test the claim that these states have the same amount of arsenic. Based on the difference, we can conclude that they do not...
                          diff 
California-Arkansas -0.7666667 
Texas-Arkansas       1.4916667  
Texas-California     2.2583333  

# Question 7
```{r, out.width="60%"}

```

# Question 8
```{r, out.width="60%"}

```

# Question 9
```{r, out.width="60%"}

```
























